### 2020-01-31 wuhan.tex
push

\newcommand{\eref}[1]{(\ref{eq:#1})}
\newcommand{\fref}[1]{Fig.~\ref{fig:#1}}
\newcommand{\Fref}[1]{Fig.~\ref{fig:#1}}
\newcommand{\sref}[1]{Sec.~\ref{#1}}
\newcommand{\frange}[2]{Fig.~\ref{fig:#1}--\ref{fig:#2}}
\newcommand{\tref}[1]{Table~\ref{tab:#1}}
\newcommand{\tlab}[1]{\label{tab:#1}}
\newcommand{\seminar}{SE\mbox{$^m$}I\mbox{$^n$}R}

@@

Since December 2019, the novel coronavirus (2019-nCoV) has been spreading in China and other parts of the world.
As of January 28th, 2020, the World Health Organization has confirmed 4593 cases, including 56 confirmed cases in 14 different countries, outside China \citep{who28report}.
Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, 
there is clear evidence that it can be directly transmitted between humans,
posing a greater threat for its spread \citep{huang2020clinical,who26report}.

@@

%% \jd{The pictures with linear scale for Ro (and r) are fine, and the Ro ones look similar. But I would say that it's a good principle to use log scales for things like Ro and r, where it's more or less clear that the important distances are multiplicative.}
%% \jd{I have no problem doing this in the next round, though. Time is important.}
\fref{eff} shows how propagating uncertainty ($\mu_r$, $\mu_G$, and $\mu_\kappa$) in different combinations would affect estimates and CIs for \Ro. For illustrative purposes, we use our pooled estimates, which represent a reasonable proxy for the state of knowledge as of 26 January. 
Comparing the models that include only some sources of uncertainty to the confidences intervals from the ``all'' model, we see that propagating error from the growth rate (which all but one of the models did) is absolutely crucial: the middle bar, which lacks growth-rate uncertainty is far too narrow.
%% \jd{I really dislike these in-text numbers, plus it misses the point that the cruciality of them depends primarily on something else (the comparison of things that leave them out with the last two bars). Who wanted this and why? I can write something about the comparison above instead?}.
%% uncertainty on \Ro, measured with our 95\% CI, is more than two fold in this case (95\% CI: 2.22--5.16).
Propagating error from the generation interval also has important effects. 
%% however (95\% CI: 2.59--4.12).
Once these two are included, the impact of leaving out uncertainty in the dispersion is small, though noticeable, in this particular example.
%% -- propagating errors from $r$ and $\bar G$ gives 95\% CI of 2.16--5.58 whereas propagating all errors gives 95\% CI of 2.06--5.89 (compare \textbf{growth rate + GI mean} with \textbf{all} in \fref{eff}).

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figure2.pdf}
\caption{
\textbf{Effects of $r$, $\bar G$, and $\kappa$ on the estimates of \Ro.}
We compare estimates of \Ro under five scenarios that propagate different combinations of uncertainties.
\textbf{base}: \Ro estimates based on the median estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$.
\textbf{growth rate}: \Ro estimates based on the the posterior distribution of $\mu_r$ while using median estimates of $\mu_G$ and $\mu_\kappa$.
\textbf{GI mean}: \Ro estimates based on the the posterior distribution of $\mu_G$ while using median estimates of $\mu_r$ and $\mu_\kappa$.
\textbf{growth rate + GI mean}: \Ro estimates based on the the joint posterior distributions of $\mu_r$ and $\mu_G$ while using a median estimate of $\mu_\kappa$.
\textbf{all}: \Ro estimates based on the joint posterior distributions of  $\mu_r$, $\mu_G$, and $\mu_\kappa$.
Vertical lines represent the 95\% confidence intervals.
%% \jsw{why not show all combinations}
%% {We discussed this, but came down on the side of simplicity. There are no examples I can think of where people included uncertainty in κ but left out one of the two big ones.}
}
\label{fig:eff}
\end{figure}

@@

Consistent with our previous observations (\fref{eff}),
we find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro (\fref{R0}).
For example, recalculating \Ro for Study 6 by using our pooled estimate of $r$ gives $\mathcal R_0 = 3.9$ (95\% CI: 2.3--9.8), which is much wider than the range they reported (2.0--3.1).
There are two explanations for this result.
First, even though the exponential growth rate $r$ and the mean generation time $\bar G$ have identical effects on \Ro under the gamma approximation framework \eref{gamma},
$r$ has a greater overall effect on \Ro because it is associated with more uncertainty;
the range of the 95\% confidence interval on $\mu_r$ spans over more than a two-fold difference, whereas the range of the 95\% confidence intervals on $\mu_G$ does not.
Second, assuming a fixed generation time ($\kappa=0$) makes the estimate of \Ro too sensitive to $r$ and $\bar G$ as explained previously.

@@

We reviewed six modeling studies of the novel coronavirus outbreak that were published online between January 24th, 2020 and January 26th, 2020 (Table 1).
Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were upload on preprint servers (bioRxiv, medRxiv, and SSRN), and one report was published from Imperial College London \citep{imaincov}.
There is a wide variation in their statistical methods and the amount of data they used to infer \Ro.
\cite{imaincov} and \cite{riouncov} simulated branching process models and compared the predicted number of cases from their models with the estimated number of total cases by January 18th.
\cite{readncov} fitted a deterministic, metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model to incidence data between January 1st and January 21st from major cities in China and other countries.
\cite{zhaoncov} and \cite{liuncov} fitted exponential growth models to incidence data up to January 22nd and January 23rd, respectively, and inferred \Ro\ via the Euler-Lotka equation \eref{euler}.
\cite{majumderncov} fitted the Incidence Decay and Exponential Adjustment (IDEA) model \citep{fisman2013idea} to incidence data up to January 26th, which is equivalent to fitting an exponential growth model and assuming a fixed generation-interval distribution.

>>>>

\newcommand{\eref}[1]{Eq.~\ref{eq:#1}}
\newcommand{\fref}[1]{Fig.~\ref{fig:#1}}
\newcommand{\Fref}[1]{Fig.~\ref{fig:#1}}
\newcommand{\sref}[1]{Sec.~\ref{#1}}
\newcommand{\frange}[2]{Fig.~\ref{fig:#1}--\ref{fig:#2}}
\newcommand{\tref}[1]{Table~\ref{tab:#1}}
\newcommand{\tlab}[1]{\label{tab:#1}}
\newcommand{\seminar}{SE\mbox{$^m$}I\mbox{$^n$}R}

@@

Since December 2019, the novel coronavirus (2019-nCoV) has been spreading in China and other parts of the world \citep{pneumonia}.
Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, its ability to directly transmit between humans has posed a greater threat for its spread \citep{huang2020clinical,who26report}.
As of January 30th, 2020, the World Health Organization (WHO) has confirmed 7818 cases, including 82 confirmed cases in 18 different countries, outside China \citep{who28report}.
WHO has now declared the outbreak a Public Health Emergency of International Concern \citep{whoemer}.

@@

\jd{The pictures with linear scale for Ro (and r) are fine, and the Ro ones look similar. But I would say that it's a good principle to use log scales for things like Ro and r, where it's more or less clear that the important distances are multiplicative.}
\jd{I have no problem doing this in the next round, though. Time is important.}
\fref{eff} shows how propagating uncertainty ($\mu_r$, $\mu_G$, and $\mu_\kappa$) in different combinations would affect estimates and CIs for \Ro. For illustrative purposes, we use our pooled estimates, which represent a reasonable proxy for the state of knowledge as of 26 January. 
Propagating error from the growth rate (which all but one of the models did) is absolutely crucial:
\jd{I really dislike these in-text numbers, plus it misses the point that the cruciality of them depends primarily on something else (the comparison of things that leave them out with the last two bars). Who wanted this and why? I can write something about the comparison above instead?}.
uncertainty in \Ro, measured with our 95\% CI, is more than two-fold in this case (95\% CI: 2.22--5.16).
Propagating error from the generation interval also has important effects, however (95\% CI: 2.59--4.12).
Once these two are included, the impact of leaving out uncertainty in the dispersion would be small, in this particular example -- 
propagating errors from $r$ and $\bar G$ gives 95\% CI of 2.16--5.58 whereas propagating all errors gives 95\% CI of 2.06--5.89 (compare \textbf{growth rate + GI mean} with \textbf{all} in \fref{eff}).

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figure2.pdf}
\caption{
\textbf{Effects of $r$, $\bar G$, and $\kappa$ on the estimates of \Ro.}
We compare estimates of \Ro under five scenarios that propagate different combinations of uncertainties.
\textbf{base}: \Ro estimates based on the median estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$.
\textbf{growth rate}: \Ro estimates based on the the posterior distribution of $\mu_r$ while using median estimates of $\mu_G$ and $\mu_\kappa$.
\textbf{GI mean}: \Ro estimates based on the the posterior distribution of $\mu_G$ while using median estimates of $\mu_r$ and $\mu_\kappa$.
\textbf{growth rate + GI mean}: \Ro estimates based on the the joint posterior distributions of $\mu_r$ and $\mu_G$ while using a median estimate of $\mu_\kappa$.
\textbf{all}: \Ro estimates based on the joint posterior distributions of  $\mu_r$, $\mu_G$, and $\mu_\kappa$.
Vertical lines represent the 95\% confidence intervals.
\jsw{why not show all combinations}
}
\label{fig:eff}
\end{figure}

@@

Consistent with our previous observations (\fref{eff}),
we find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro (\fref{R0}).
For example, recalculating \Ro for Study 6 by using our pooled estimate of $r$ gives $\mathcal R_0 = 3.9$ (95\% CI: 2.3--9.8), which is much wider than the range they reported (2.0--3.1).
There are two explanations for this result.
First, even though the exponential growth rate $r$ and the mean generation time $\bar G$ have identical effects on \Ro under the gamma approximation framework (\eref{gamma} in Methods),
$r$ has a greater overall effect on \Ro because it is associated with more uncertainty;
the range of the 95\% confidence interval on $\mu_r$ spans over more than a two-fold difference, whereas the range of the 95\% confidence intervals on $\mu_G$ does not.
Second, assuming a fixed generation time ($\kappa=0$) makes the estimate of \Ro too sensitive to $r$ and $\bar G$ as explained previously.

@@

We reviewed six modeling studies of the novel coronavirus outbreak that were published online between January 24th, 2020 and January 26th, 2020 (Table 1).
Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were upload on preprint servers (bioRxiv, medRxiv, and SSRN), and one report was published from Imperial College London \citep{imaincov}.
There is a wide variation in their statistical methods and the amount of data they used to infer \Ro.
\cite{imaincov} and \cite{riouncov} simulated branching process models and compared the predicted number of cases from their models with the estimated number of total cases by January 18th.
\cite{readncov} fitted a deterministic, metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model to incidence data between January 1st and January 21st from major cities in China and other countries.
\cite{zhaoncov} and \cite{liuncov} fitted exponential growth models to incidence data up to January 22nd and January 23rd, respectively, and inferred \Ro\ via the Euler-Lotka equation (\eref{euler}).
\cite{majumderncov} fitted the Incidence Decay and Exponential Adjustment (IDEA) model \citep{fisman2013idea} to incidence data up to January 26th, which is equivalent to fitting an exponential growth model and assuming a fixed generation-interval distribution.
